---
title: "Take-home_Ex03"
title-block-banner: true
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
editor: visual
author: Yen Yun Hsuan
date: "11 June 2023"
execute: 
  warning: false
---

## Background

This is [Mini-Challenge 3](https://vast-challenge.github.io/2023/MC3.html) of VAST Challenge 2023.

FishEye International, a non-profit focused on countering illegal, unreported, and unregulated (IUU) fishing, has been given access to an international finance corporation\'s database on fishing related companies. FishEye has transformed the database into a knowledge graph. It includes information about companies, owners, workers, and financial status. FishEye is aiming to use this graph to **identify anomalies** that could indicate a company is involved in IUU.

FishEye analysts have attempted to use traditional node-link visualizations and standard graph analyses, but these were found to be ineffective because the scale and detail in the data can obscure a business\'s true structure. Can you help FishEye develop a **new visual analytics approach** to better understand fishing business anomalies?

## Tasks and Questions:

Develop a visual analytics process to find **similar businesses and group the**m. This analysis should focus on a business\'s **most important features and present** those features clearly to the user. Limit your response to 400 words and 5 images.

Data preparation

## Installing and launching R packages

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts,jsonlite,gtsummary,
               tinygraph)
```

## Import data

```{r}
MC3 <- fromJSON("data/MC3.json")
```

```{r}
MC3_nodes <- as_tibble(MC3$nodes) 

glimpse(MC3_nodes)
```

```{r}
MC3_edges <- as_tibble(MC3$links) 
 
glimpse(MC3_edges)
```
