---
title: "Take Home Exercise 1"
title-block-banner: true
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
editor: visual
author: Yen Yun Hsuan
date: "13 May 2023"
execute: 
  warning: false
---

# Task

The local council of a city is in the process of preparing the Local Plan 2023. A **sample survey of 1000 representative residents** had been conducted to collect data related to their household demographic and spending patterns. The city aims to use the data to assist with their **major community revitalization efforts**, including how to allocate renewal grant.

This take-home exercise are required to reveal the **demographic and financial characteristics** of the city by using statistical graphics methods.

# Data Preparation

## Import package and data

```{r}
pacman::p_load(patchwork, tidyverse, ggstatsplot, 
               ggdist, gganimate, png, gifski, nortest,dplyr,tidyr,lubridate,skimr,ggcorrplot,ggpubr,plotly,
               ggiraph, DT,ggridges,viridis )
```

```{r}
participant <- read_csv("data/Participants.csv")
financial <- read_csv("data/FinancialJournal.csv")
```

Two data sets are provided. They are:

## Participants

Contains information about the residents of City of Engagement that have agreed to participate in this study.

### Change data format

Change dbl format (household size, age, participant ID) to **int (integer)** since these values should not be float.

```{r}
participant <- participant %>% 
mutate_at(vars(householdSize,age,participantId), list(~as.integer(.))) 
```

Reorder education level from lowest to highest degree.

```{r}
participant <- participant %>%
  mutate(educationLevel=factor(educationLevel,levels = c("Low", "HighSchoolOrCollege", "Bachelors", "Graduate")))
data.frame(levels(participant$educationLevel))
```

Each column format has been adjusted as below table shown.

```{r}
head(participant)
```

## Financial Journal

Contains information about financial transactions.

### Create Year-Month column

Use lubridate package (ymd_hms) to transform timestamp column into datetime format column "time". Create new columns "**Year_Month**" and "Year", since time information is no need for following visual analysis.

```{r}
#| code-fold: true
#| code-summary: "Show code"
financial$time <- as.Date(ymd_hms(financial$timestamp))
financial <- financial %>% 
  mutate(Year_Month = format(financial$time, "%Y-%m"),
         Year_Month = factor((Year_Month),levels = unique(Year_Month)),
         Year = format(financial$time, "%Y"),
         participantId = as.integer(participantId))
head(financial)

```

### Group transaction by category

There are some duplicate records with same participant ID, category and timestamp. Use distinct function to **drop duplicate rows.** The original amount column contains positive(meaning income) and negative (expense) values. Transform amount to **absolute value** to avoid confusion.

```{r}
#| code-fold: true
#| code-summary: "Show code"
financial_new <- 
  financial %>% 
  distinct(participantId, category, timestamp,.keep_all = TRUE) %>%
  group_by(participantId, category, Year,Year_Month) %>%
  summarise(amount = round(sum(abs(amount)),1), .groups = 'drop') 

```

To compare amount of different categories, need to split out the unique values in the "category" column and their corresponding amounts into new columns by [pivot_wider](https://tidyr.tidyverse.org/reference/pivot_wider.html) function. Fill null values with 0.

```{r}
#| code-fold: true
#| code-summary: "Show code"
financial_orginize <- financial_new %>% 
  pivot_wider(names_from = category, values_from = amount, values_fill = 0) %>% 
  mutate(Total_Expense=rowSums(select(.,Education,Food,Recreation,Shelter)))

head(financial_orginize)
```

## Merge two datasets

Some participants only have 1-month records, so remain participants' records with 12 months, and calculate their monthly average wage and spending in the "financial_average" data frame.

```{r}
#| code-fold: true
#| code-summary: "Show code"
financial_average <- financial_orginize %>% 
  group_by(participantId) %>%
  summarise(across(Education:Total_Expense,sum), 
            Month_count = n_distinct(Year_Month), 
            .groups = 'drop') %>%
  filter(Month_count == 12) %>% 
  mutate(across(Education:Total_Expense, ~round(. / Month_count, 1)))


```

Join with "participant" by the same participant ID.

```{r}
df_average <-  merge(x=financial_average,y=participant,by="participantId")

DT::datatable(df_average)

```

# Visualization

## Correlation with joviality

### What is key factor to joviality?

To better understand the correlation between joviality, annual income and spending, comparing the Pearson correlation values. Major findings as follow:

1.  Higher income does not necessarily lead to higher joviality

    With closer examination of the correlation scatter plots, can find that there are participants with extremely high wage. While the majority of participants with wages below the 95th percentile exhibit a wide range of joviality values, those in the **top 5th percentile** of wage earners tend to have joviality values below 0.5.

2.  Joviality is most correlated with recreation spending.

    For participants with spending above the 22th percentile, can observe a positive correlation between spending on recreation and joviality. This suggests that for those who spend more on recreation activities, there is a tendency to experience higher levels of joviality.

::: panel-tabset
## Correlation

```{r}
ggstatsplot::ggcorrmat(
  data = df_average,
  cor.vars = c("Wage","joviality","Recreation","Food","Shelter"))
```

## Wage v.s Joviality

```{r}
ggscatterstats(
  data = df_average,
  x = Wage,
  y = joviality,
  title = 'Wage vs Jovality',
  ggplot.component = list(
    scale_x_continuous(
      breaks = seq(0, 15000, 2000),
      limits = c(0, 15000)
    )
  )
) +
  geom_vline(aes(xintercept = quantile(Wage,probs=0.95)), color = "darkgreen",linetype="dashed",linewidth=1)+
  geom_hline(aes(yintercept = 0.5), color = "darkorange",linetype="dashed",linewidth=1)+
  annotate(
    "text",
    x = quantile(df_average$Wage, probs = 0.95),
    y = -0.05,
    label = "95th Percentile",
    color = "darkgreen"
  )

```

## Recreation expense v.s Joviality

```{r}
ggscatterstats(
  data = df_average,
  x = Recreation,
  y = joviality,
  title = 'Recreation vs Jovality',
  ggplot.component = list(
    scale_x_continuous(
      breaks = seq(0, 900, 100),
      limits = c(0, 900)
    )
  )
) +
  geom_vline(aes(xintercept = quantile(Recreation,probs=0.22)), color = "darkgreen",linetype="dashed",linewidth=1)+
  #geom_hline(aes(yintercept = 0.5), color = "darkorange",linetype="dashed",linewidth=1)+
  annotate(
    "text",
    x = quantile(df_average$Recreation, probs = 0.22),
    y = -0.05,
    label = "22th Percentile",
    color = "darkgreen"
  )
```
:::

## Financial pattern

### What is overall transaction distribution and difference

Since each category has a different value range, applying scale of log10 to make the ranges comparable in the same graph. Education expense has the lowest average value, while recreation expense exhibits a larger variance, and food expense is more centered.

```{r}
# Reshape the data from wide to long format using pivot_longer
p <- df_average %>%
  group_by(participantId) %>%
  summarise(across(Education:Wage),
            .groups = 'drop') %>% 
  pivot_longer(cols = -participantId, names_to = "Category", values_to = "Value")

plot <- ggplot(p,aes(x = Value, fill = Category)) +
  geom_density_interactive(aes(tooltip=Category),alpha=0.8) +
# Generate a discrete version from the "viridis" color palette to make the colors more beautiful
  scale_fill_viridis_d(option = "D")+
  scale_y_continuous(NULL,           
                     breaks = NULL)+
  ggtitle("Financial transaction distribution")+
  scale_x_log10()

# Create girafe interactive label to easily identify each density category
girafe(
  ggobj = plot,
  width_svg = 6,
  height_svg = 6 * 0.618
)                        

```

Upon closer examination of the expenses in the following categories, food and recreation have similar mean values, while recreation exhibits a larger spread of spending amounts. And shelter expense demonstrates the highest number of outliers.

```{r}
# Use same color from previous plot for the same category
color_palette <- viridis_pal(option = "D")(5)
selected_colors <- color_palette[2:4] 

p1 <- df_average %>%
  group_by(participantId) %>%
  summarise(across(Food:Shelter),
            .groups = 'drop') %>% 
  pivot_longer(cols = -participantId, names_to = "Category", 
               values_to = "Value") %>%
  ggplot(aes(y = Value, x = Category, fill = Category)) +
  geom_boxplot() +
  scale_fill_manual(values = selected_colors)+
  stat_summary(geom = "point",       
               fun.y = "mean",         
               color = "black",  
               size = 4) +
  labs(x = "Spending category", y = "Amount") +
  theme(legend.position = "none")

ggplotly(p1)

```

### Standard error bars of mean

To assess the standard error of spending and wage amounts, use z-score normalization to standardize the range. Wage exhibits the largest range of error bars, suggesting a significant level of variability and uncertainty surrounding the wage data from these participants.

::: panel-tabset
## Data

```{r}
p <- df_average %>%
  group_by(participantId) %>%
  summarise(across(Education:Wage),
            .groups = 'drop')
# Select all columns except for participant id, and apply mean/sd calculation to all selected columns
means <- apply(p[, -1], 2, mean)
std_errors <- apply(p[, -1], 2, function(x) sd(x) / sqrt(length(x)-1))

# Create a new dataframe with means and standard errors by category
plot_df <- data.frame(
  variable = names(means),
  mean = means,
  se = std_errors
)

plot_df <- rownames_to_column(plot_df, var = "row_index") %>%
  select(-row_index)

knitr::kable(head(plot_df), format = 'html')
```

## Plot

```{r}
plot_df$normalized_mean <- scale(plot_df$mean)

ggplot(plot_df) +
  geom_errorbar_interactive(aes(x=variable,ymin = normalized_mean - se, ymax = normalized_mean + se), width = 0.2) +
  geom_point(aes
           (x=variable,y=normalized_mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1)+
  labs(title = "Normalized Mean and Standard Error", x = "Variable", y = "Mean") 

```
:::

## With/ without kids

```{r}
#| code-fold: true
#| code-summary: "Show code"
p <- ggplot(data=df_average, 
       aes(x=haveKids,fill=educationLevel)) +
  geom_bar()+
# +
#   geom_text(stat="count",
#       aes(label=paste0(
#       "(",count,") ",round(..count../sum(..count..)*100, 1), "%")),
#       vjust=-1) +
  ylab("Count") +
  xlab("")+
  coord_cartesian(ylim=c(0,800))
  theme(axis.title.y=element_text(angle = 0))
ggplotly(p)
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
summary_kid <- df_average %>%
  group_by(haveKids) %>%
  summarise(across(Education:RentAdjustment,sum),
            persons = n_distinct(participantId),
            .groups = 'drop') %>%
  mutate(across(Education:RentAdjustment, ~round(. / persons, 1)))

melted_kid <- gather(summary_kid, key = "Category", value = "Value", -haveKids)
melted_kid <- melted_kid %>% filter(Category != "persons",Category  
                                    !="RentAdjustment",Category != "Wage")
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
prop <- ggplot(melted_kid, 
                aes(Category, Value, fill =  haveKids)) + 
  geom_bar(stat = "identity", position = "dodge") +

  labs(title = "Spending of have/ not having kids", y = "Average Amount", fill = "Have Kids") +
  ylab("Amount") +
   theme(
        plot.title = element_text(hjust = 0.4, size = 15, face = 'bold'),
        plot.margin = margin(20, 20, 20, 20),
        legend.position = "right",
        axis.text = element_text(size = 8, face = "bold"),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1),
        axis.title.x = element_text(hjust = 0.5, size = 12),
        axis.title.y = element_text(angle=0,hjust = 0.5, size = 12))

ggplotly(prop)
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
prop <- ggplot(melted_kid, 
                aes(haveKids, Value, fill =  Category )) + 
  geom_bar(stat = "identity",position = "dodge") +

  labs(title = "Spending of have/ not having kids", y = "Average Amount", fill = "Have Kids") +
  ylab("Amount") +
   theme(
        plot.title = element_text(hjust = 0.4, size = 15, face = 'bold'),
        plot.margin = margin(20, 20, 20, 20),
        legend.position = "right",
        axis.text = element_text(size = 8, face = "bold"),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1),
        axis.title.x = element_text(hjust = 0.5, size = 12),
        axis.title.y = element_text(angle=0,hjust = 0.5, size = 12))
prop
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
ggplot(data = df_average, aes(fill = haveKids)) +
  geom_density(aes(x = Wage), linewidth = 1, alpha = 0.7) +
  ggtitle("Wage with/without Kids") +
  scale_y_continuous(NULL, breaks = NULL) +
  theme_minimal()+
  geom_vline(aes(xintercept=mean(Wage)),
             color="red", 
             linetype="dashed", 
             size=1) +
  geom_vline(aes(xintercept=median(Wage)),
             color="grey30",
             linetype="dashed", 
             size=1)+
  annotate("text", x = mean(df_average$Wage), 
           y = 0, label = "Mean", color = "red", vjust = -1) +
  annotate("text", x = median(df_average$Wage), 
           y = 0, label = "Median", color = "grey30", vjust = 1)

```

## Different education level

```{r}
ggplot(data=df_average, 
       aes(x=educationLevel)) +
  geom_bar()+
  geom_text(stat="count", 
      aes(label=paste0( 
      round(..count../sum(..count..)*100, 0), "%")),
      vjust=-1) +
  coord_cartesian(ylim=c(0,700))+
  ylab("No. of\nPartipipant") +
  theme(axis.title.y=element_text(angle = 0))
```

```{r}
ggplot(data=df_average, aes(x = Wage,fill=educationLevel)) +
  geom_density(alpha=0.5)+
  ggtitle("Distribution of Wage by education level")+      
  scale_y_continuous(NULL,           
                     breaks = NULL)+
    coord_cartesian(xlim=c(2500,20000))

```

```{r}
ggbetweenstats(
  data = df_average,
  x = educationLevel, 
  y = Wage,
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
summary <- df_average %>%
  group_by(educationLevel) %>%
  summarise(across(Education:Shelter,sum),
            persons = n_distinct(participantId),
            .groups = 'drop') %>%
  mutate(across(Education:Shelter, ~round(. / persons, 1)))

melted <- gather(summary, key = "Category", value = "Value", -educationLevel)
melted <- melted %>% filter(Category != "persons")
melted
```

```{r}
prop <- ggplot(melted, 
                aes(educationLevel , Value, fill =Category  )) + 
  geom_bar(stat = "identity") +
  labs(title = "Spending of each education level", y = "Average Amount", 
       fill = "education level") +
  ylab("Amount") +
   theme(
        plot.title = element_text(hjust = 0.4, size = 15, face = 'bold'),
        plot.margin = margin(20, 20, 20, 20),
        legend.position = "right",
        axis.text = element_text(size = 8, face = "bold"),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1),
        axis.title.x = element_text(hjust = 0.5, size = 12),
        axis.title.y = element_text(angle=0,hjust = 0.5, size = 12))

ggplotly(prop)
```

```{r}
ggplot(data=df_average, aes(y = Total_Expense,x=haveKids, fill=educationLevel)) +
  geom_boxplot()+
  ggtitle("Total Expense")+
  geom_point(position="jitter", 
             size = 0.2)
```

## Difference among age

```{r}
df_age <- df_average %>% 
  mutate(age_bins = 
           cut(age, 
               breaks = c(17,25,35,45,55,60))
)
```

```{r}
ggbetweenstats(
  data = df_age,
  x = age_bins, 
  y = Wage,
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

```{r}
# write.csv(df_average, "df_average.csv", row.names=FALSE)
# write.csv(financial_average, "financial_average.csv", row.names=FALSE)
```

```{r}
summary_age <- df_age %>%
  group_by(age_bins) %>%
  summarise(across(Education:Shelter,sum),
            persons = n_distinct(participantId),
            .groups = 'drop') %>%
  mutate(across(Education:Shelter, ~round(. / persons, 1)))

melted_age <- gather(summary_age, key = "Category", value = "Value", -age_bins)
melted_age <- melted_age %>% filter(Category != "persons")
```

```{r}
ggplot(df_age, aes(x = Total_Expense, y = age_bins,fill=age_bins)) + 
  geom_density_ridges() + 
  theme_ridges() 
```
